---
title: "Activity: Bird Species"
author: "Your Names Here"
date: "DUE DATE "
output: html_notebook
---


# Set Up

```{r}
# Load `tidyverse` and `dcData` packages




# Load the `OrdwayBirds` data set into your RStudio environment from the `dcData` package




# Show a few commands used to inspect the data set a few different ways



```

# Activity

When you're finished with this activity, you will have a graph that shows what time of year various species appear at the Katherine Ordway Natural History Study Area in Inver Grove Heights, MN.

### Step 0

Before we begin, the book gives us some instructions to select a few key variables and clean up the date formatting.  Use the commands given in the book, but be sure to study the functions used so you know what is happening.  In this case, we select `SpeciesName`, `Month`, and `Day`. Then a `mutate()` command converts each variable to character, and then converts the resulting "character string" to a number.

```{r}
OrdwayBirds <-
  OrdwayBirds %>%
  select( SpeciesName, Month, Day ) %>%
  mutate( Month = as.numeric(as.character(Month)), 
          Day = as.numeric(as.character(Day)))


```



### Step 1




```{r}
OrdwayBirds %>%
  summarise(count = n_distinct(SpeciesName))
  # count 275

OrdwaySpecNameCount <-
  OrdwaySpeciesNames %>%
  summarise(count = n_distinct(SpeciesNameCleaned))
OrdwaySpecNameCount 
# count 109


```


### Step 2
  



```{r}
Corrected <-
  OrdwayBirds %>%
  inner_join(OrdwaySpeciesNames, by = "SpeciesName") %>% 
  select(Species = SpeciesNameCleaned, Month, Day) %>%
  na.omit()  


```


### Step 3

We want to identify the top 5-6 bird species sighted at the Katherine Ordway Natural History Study Area.

```{r}
CountCorrect <- 
  Corrected %>% 
  group_by(Species) %>% 
  summarise(count=n()) %>%
  arrange(desc(count))

topSix <-
  CountCorrect %>% 
  head(n = 6) %>% 
  .$Species
topSix

Majors <-
  Corrected %>% 
  filter(Species %in% topSix)


```


### Step 4

We now want to produce a month-by-month count for each of the major species, and then construct a chart to tell the story of what time of year the various species appear.

```{r}
ByMonth <- Majors %>%
  group_by(Species, Month) %>%
  summarise(count = n()) %>%
  arrange(Month)


```

Now use the ggplot template below to generate the desired graph. (Remark you can also use esquisser or mplot to create a line chart instead of the barchart)



```{r}
CrazyMonth <- monthAbbr <- with(Majors, plyr::mapvalues(Month, from = 1:12, to = month.abb))
CrazyMonth <- factor(CrazyMonth, levels = month.abb)
Majors$Month <- CrazyMonth

# ggplot commands
Majors %>% 
  ggplot(aes(x = Month)) +
  geom_bar() +
  facet_wrap(~ Species) +
  theme(axis.text.x=element_text(angle=45,hjust=1))

```

According to the graph, we can answer the questions posed:

##### 1. Which species are present year-round?  

the American Crow and Black-capped Chickadee are present year-round

##### 2. Which species are migratory, that is, primarily present in one or two seasons?

The Lincolns Sparrow and Slate-colored Junco are primarily present in one or two seasons

##### 3. What is the peak month for each major species? (bullet list)

- American Goldfinch: October
- Black-capped Chickadee: November
- Field Sparrow: May
- Lincolns Sparrow: October
- Slate-colored Junco: October
- Tree swallow: March


##### 4. Which major species are seen in good numbers for at least 6 months of the year?

The species that are seen in good numbers for at least 6 months of the year are the Black-capped Chickadee, Red-winged Blackbird, American Crow, and Dark-eyed Junco

```{r}
n_distinct()
```

